---
title: "hannahli09"
author: "Hannah Li, 8264178"
date: "10/18/2022"
output: html_document
---


# Questions for professor
1. file urls not working
2. trimming - only ended up with 1 outlier
3. question 3- "repeat the analysis" does this mean repeat all three methods in the report or like exploratory analysis
4. what is fuzzy intersections vs hard intersection

----------------
# other things to do
5. sensitivity and accuracy tables/ visuals in question 3
6. make final table with  sensitivity and accuracy at the end 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(infer)
library(randomForest)
library(tidymodels)
library(modelr)
library(yardstick)
```

```{r}
View(biomarker_clean)
```

Question 2:
Temporarily remove the outlier trimming from preprocessing and do some exploratory analysis of outlying values. 
Are there specific subjects (not values) that seem to be outliers? If so, are outliers more frequent in one 
group or the other? (Hint: consider tabluating the number of outlying values per subject.)

```{r}
head(biomarker_clean)
```

from the preprocessing file
```{r}
# url <- 'https://raw.githubusercontent.com/pstat197/pstat197a/main/materials/labs/lab3-iteration/data/biomarker-raw.csv'


# file path for Kassandra: /Users/kassandratrejo/Desktop/PSTAT197A/biomarkers-group-3/data/biomarker-raw.csv

var_names <- read_csv('/Users/hannahli/Documents/Github/biomarkers-group-3/data/biomarker-raw.csv', 
                     col_names = F, 
                     n_max = 2, 
                     col_select = -(1:2)) %>%
  t() %>%
  as_tibble() %>%
  rename(name = V1, 
         abbreviation = V2) %>%
  na.omit()
```


```{r}
# function for trimming outliers (good idea??)
trim <- function(x, .at){
  x[abs(x) > .at] <- sign(x[abs(x) > .at])*.at
  return(x)
}
```

remove the trimming 
```{r}
# file path for Kassandra: /Users/kassandratrejo/Desktop/PSTAT197A/biomarkers-group-3/data/biomarker-raw.csv

biomarker_clean2 <- read_csv('/Users/hannahli/Documents/Github/biomarkers-group-3/data/biomarker-raw.csv', 
         skip = 2,
         col_select = -2L,
         col_names = c('group', 
                       'empty',
                       pull(var_names, abbreviation),
                       'ados'),
         na = c('-', '')) %>%
  filter(!is.na(group)) %>%
  # log transform, center and scale, and trim
  mutate(across(.cols = -c(group, ados), 
                ~ scale(log10(.x))[, 1], .at = 3)) %>%
  # reorder columns
  select(group, ados, everything())
```
 
```{r}
view(biomarker_clean2)
```

```{r}
biomarker_clean <- read_csv('/Users/hannahli/Documents/Github/biomarkers-group-3/data/biomarker-raw.csv', 
         skip = 2,
         col_select = -2L,
         col_names = c('group', 
                       'empty',
                       pull(var_names, abbreviation),
                       'ados'),
         na = c('-', '')) %>%
  filter(!is.na(group)) %>%
  # log transform, center and scale, and trim
  mutate(across(.cols = -c(group, ados), 
                ~ trim(scale(log10(.x))[, 1], .at = 3))) %>%
  # reorder columns
  select(group, ados, everything())


boxplot(biomarker_clean[,2]) #ados

boxplot(biomarker_clean2[,3:50]) # all proteins

ggplot(biomarker_clean2) +
  aes(x = "", y = group) +
  geom_boxplot(fill = "#0c4c8a") +
  theme_minimal()



biomarker_clean2 %>%
  group_by(group)%>%
  summarise(n=mean)

install.packages('dlookr')
library(dlookr)
outliers <- diagnose_outlier(biomarker_clean2)
outliers %>%
  arrange(desc(outliers_cnt))
```

Experiment with the training data 

1. repeat the analysis but carry out the entire selection procedure on a training partition.- set aside some testing data for the end 
2. choose a larger number of top predictive proteins using each selection method (USE 20 PROTEINS)
3. use a fuzzy intersection instead of a hard intersection to combine
the sets of top predictive proteins across selection methods. 

```{r}
set.seed(638)
biomarker_split<-initial_split(biomarker_clean, prop= 0.70,strata=group)
biomarker_train<-training(biomarker_split)
biomarker_test<-testing(biomarker_split)
```

## Random Forest method
```{r}
predictors<- biomarker_train %>% 
  select(-c(group,ados))

response <- biomarker_train %>%  pull(group) %>%  factor()

# fit the Random Forest 
random_f<-randomForest(x= predictors, 
                       y = response,
                       ntree= 2000,
                       importance= T)
# check the errors 
random_f$confusion

# compute the importance scores 
proteins_1<-random_f$importance %>% 
  as_tibble() %>% 
  mutate(protein= rownames(random_f$importance)) %>% 
  slice_max(MeanDecreaseGini, n = 20) %>% 
  pull(protein)
```

```{r}
proteins_1 # top 20
```
## T-test method

- t-test determines if there is a significant difference between the means of two groups.
choose the top proteins with the most significant t-test values to use in the prediction model


## Correlation method

- measures the statistical relationship between two variables (calculate each protein's correlation with ADOS total scores as a measure of ASD severity). 
choose the top proteins with the highest absolute values of the correlation coefficient to use.


## Logistic regression model 
- used with datasets based on the random forest, t-test, and correlation models (group (ASD or TD) is output)
- randomly assigned 80% as training and 20% as test data. 
- calculated model's AUC to evaluate it.