---
title: "hannahli09"
author: "Hannah Li, 8264178"
date: "10/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(infer)
library(randomForest)
library(tidymodels)
library(modelr)
library(yardstick)
```

```{r}
View(biomarker_clean)
```

Question 2:
Temporarily remove the outlier trimming from preprocessing and do some exploratory analysis of outlying values. 
Are there specific subjects (not values) that seem to be outliers? If so, are outliers more frequent in one 
group or the other? (Hint: consider tabluating the number of outlying values per subject.)

```{r}
head(biomarker_clean)
```

from the preprocessing file
```{r}
var_names <- read_csv('data/biomarker-raw.csv', 
                     col_names = F, 
                     n_max = 2, 
                     col_select = -(1:2)) %>%
  t() %>%
  as_tibble() %>%
  rename(name = V1, 
         abbreviation = V2) %>%
  na.omit()
```


```{r}
# function for trimming outliers (good idea??)
trim <- function(x, .at){
  x[abs(x) > .at] <- sign(x[abs(x) > .at])*.at
  return(x)
}
```

remove the trimming 
```{r}
biomarker_clean2 <- read_csv('/Users/kassandratrejo/Desktop/PSTAT197A/biomarkers-group-3/data/biomarker-raw.csv', 
         skip = 2, 
         col_select = -2L,
         col_names = c('group', 
                       'empty',
                       pull(var_names, abbreviation),
                       'ados'),
         na = c('-', '')) %>%
  filter(!is.na(group)) %>%
  # log transform, center and scale, and trim
   #mutate(across(.cols = -c(group, ados), 
               # ~ trim(scale(log10(.x))[, 1], .at = 3))) %>%
  # reorder columns
  select(group, ados, everything())
```

```{r}
view(biomarker_clean2)
```

```{r}
biomarker_clean2 <- read_csv('/Users/kassandratrejo/Desktop/PSTAT197A/biomarkers-group-3/data/biomarker-raw.csv', 
         skip = 2, 
         col_select = -2L,
         col_names = c('group', 
                       'empty',
                       pull(var_names, abbreviation),
                       'ados'),
         na = c('-', '')) %>% 
  select(group, ados, everything())
```

