---
title: "hannahli09"
author: "Hannah Li, 8264178"
date: "10/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(infer)
library(randomForest)
library(tidymodels)
library(modelr)
library(yardstick)
```

```{r}
View(biomarker_clean)
```

Question 2:
Temporarily remove the outlier trimming from preprocessing and do some exploratory analysis of outlying values. 
Are there specific subjects (not values) that seem to be outliers? If so, are outliers more frequent in one 
group or the other? (Hint: consider tabluating the number of outlying values per subject.)

```{r}
head(biomarker_clean)
```

from the preprocessing file
```{r}
var_names <- read_csv('/Users/kassandratrejo/Desktop/PSTAT197A/biomarkers-group-3/data/biomarker-raw.csv',
                     col_names = F, 
                     n_max = 2, 
                     col_select = -(1:2)) %>%
  t() %>%
  as_tibble() %>%
  rename(name = V1, 
         abbreviation = V2) %>%
  na.omit()
```


```{r}
# function for trimming outliers (good idea??)
trim <- function(x, .at){
  x[abs(x) > .at] <- sign(x[abs(x) > .at])*.at
  return(x)
}
```

remove the trimming 
```{r}
biomarker_clean2 <- read_csv('/Users/kassandratrejo/Desktop/PSTAT197A/biomarkers-group-3/data/biomarker-raw.csv', 
         skip = 2, 
         col_select = -2L,
         col_names = c('group', 
                       'empty',
                       pull(var_names, abbreviation),
                       'ados'),
         na = c('-', '')) %>%
  filter(!is.na(group)) %>%
  # log transform, center and scale, and trim
  mutate(across(.cols = -c(group, ados), 
                ~ scale(log10(.x)[,1]))) %>%
  # reorder columns
  select(group, ados, everything())
```

```{r}
view(biomarker_clean2)
```

```{r}
biomarker_clean2 <- read_csv('/Users/kassandratrejo/Desktop/PSTAT197A/biomarkers-group-3/data/biomarker-raw.csv', 
         skip = 2, 
         col_select = -2L,
         col_names = c('group', 
                       'empty',
                       pull(var_names, abbreviation),
                       'ados'),
         na = c('-', '')) %>% 
  select(group, ados, everything())
```

Experiment with the training data 

1. repeat the analysis but carry out the entire selection procedure on a training partition.- set aside some testing data for the end 
2. choose a larger number of top predictive proteins using each selection method 
3. use a fuzzy intersection instead of a hard intersection to combine
the sets of top predictive proteins across selection methods. 

```{r}
set.seed(638)
biomarker_split<-initial_split(biomarker_clean, prop= 0.70,strata=group)
biomarker_train<-training(biomarker_split)
biomarker_test<-testing(biomarker_split)
```

Random Forest
```{r}
predictors<- biomarker_train %>% 
  select(-c(group,ados))

response <- biomarker_train %>%  pull(group) %>%  factor()

# fit the Random Forest 
random_f<-randomForest(x= predictors, 
                       y = response,
                       ntree= 2000,
                       importance= T)
# check the errors 
random_f$confusion

# compute the importance scores 
proteins_1<-random_f$importance %>% 
  as_tibble() %>% 
  mutate(protein= rownames(random_f$importance)) %>% 
  slice_max(MeanDecreaseGini, n = 20) %>% 
  pull(protein)
```

```{r}
proteins_1
```

