---
title: "Question 3"
output: html_document
date: "2022-10-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r} 
library(tidyverse)
library(infer)
library(randomForest)
library(tidymodels)
library(modelr)
library(yardstick)
```

```{r}
View(biomarker_clean)
```


Experiment with the training data 

1. repeat the analysis but carry out the entire selection procedure on a training partition.- set aside some testing data for the end 
2. choose a larger number of top predictive proteins using each selection method 
3. use a fuzzy intersection instead of a hard intersection to combine
the sets of top predictive proteins across selection methods. 

```{r}
set.seed(638)
biomarker_split<-initial_split(biomarker_clean, prop= 0.70,strata=group)
biomarker_train<-training(biomarker_split)
biomarker_test<-testing(biomarker_split)
```

Random Forest
```{r}
predictors<- biomarker_train %>% 
  select(-c(group,ados))

response <- biomarker_train %>%  pull(group) %>%  factor()

# fit the Random Forest 
random_f<-randomForest(x= predictors, 
                       y = response,
                       ntree= 2000,
                       importance= T)
# check the errors 
random_f$confusion

# compute the importance scores 
proteins_1<-random_f$importance %>% 
  as_tibble() %>% 
  mutate(protein= rownames(random_f$importance)) %>% 
  slice_max(MeanDecreaseGini, n = 20) %>% 
  pull(protein)
```

```{r}
proteins_1
```

T-testing 
```{r}
tt_fn<-function(.df) {
  t_test(.df,
         formula = level ~ group,
         order= c('ASD', 'TD'), alternative = 'two-sided',
         var.equal = F)
}

ttests_1<- biomarker_train %>% 
  select(-ados) %>% 
  pivot_longer(-group, names_to = 'protein', values_to = 'level' ) %>% 
  nest(data=c(level, group)) %>% 
  mutate(ttest = map(data, tt_fn)) %>% 
  unnest(ttest) %>% 
  arrange(p_value) %>% 
  mutate( r= n(), hm= log(r) +1/(2*r) - digamma(1), rank=row_number(), 
          p.adj= r*hm*p_value/rank)

proteins_2<-ttests_1 %>% 
  slice_min(p.adj, n =20) %>% 
  pull(protein)
```

```{r}
proteins_2
```

Logistic Regression 

```{r}
proteins_subset<-intersect(proteins_1,proteins_2)

biomarker_sset<- biomarker_train %>% 
  select(group, any_of(proteins_subset)) %>% 
  mutate(class= (group == 'ASD')) %>% 
  select(-group)

set.seed(549)
biomarker_split2<-biomarker_sset %>% 
  initial_split(prop=0.70)

fit<-glm(class ~., data=training(biomarker_split2), family= 'binomial')

class_metrics<-metric_set(sensitivity, specificity, accuracy,roc_auc)

testing(biomarker_split2) %>% 
  add_predictions(fit, type= 'response') %>% 
  class_metrics(estimate = factor(pred > 0.5), 
                truth= factor(class), pred, event_level = 'second')
```

top 10 
```{r}
fit %>% 
  broom::tidy() %>% 
  knitr::kable() 
```

