---
title: "Question 3"
output: html_document
date: "2022-10-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r} 
library(tidyverse)
library(infer)
library(randomForest)
library(tidymodels)
library(modelr)
library(yardstick)
```

```{r}
View(biomarker_clean)
```


Experiment with the training data 

1. repeat the analysis but carry out the entire selection procedure on a training partition.- set aside some testing data for the end 
2. choose a larger number of top predictive proteins using each selection method (USE 20 PROTEINS)
3. use a fuzzy intersection instead of a hard intersection to combine
the sets of top predictive proteins across selection methods. 

```{r}
set.seed(638)
biomarker_split<-initial_split(biomarker_clean, prop= 0.70,strata=group)
biomarker_train<-training(biomarker_split)
biomarker_test<-testing(biomarker_split)
```

Random Forest
```{r}
predictors<- biomarker_train %>% 
  select(-c(group,ados))

response <- biomarker_train %>%  pull(group) %>%  factor()

# fit the Random Forest 
random_f<-randomForest(x= predictors, 
                       y = response,
                       ntree= 2000,
                       importance= T)
# check the errors 
random_f$confusion

# compute the importance scores 
proteins_1<-random_f$importance %>% 
  as_tibble() %>% 
  mutate(protein= rownames(random_f$importance)) %>% 
  slice_max(MeanDecreaseGini, n = 20) %>% 
  pull(protein)
```

```{r}
proteins_1 # top 20
```

T-testing 
```{r}
tt_fn<-function(.df) {
  t_test(.df,
         formula = level ~ group,
         order= c('ASD', 'TD'), alternative = 'two-sided',
         var.equal = F)
}

ttests_1<- biomarker_train %>% 
  select(-ados) %>% 
  pivot_longer(-group, names_to = 'protein', values_to = 'level' ) %>% 
  nest(data=c(level, group)) %>% 
  mutate(ttest = map(data, tt_fn)) %>% 
  unnest(ttest) %>% 
  arrange(p_value) %>% 
  mutate( r= n(), hm= log(r) +1/(2*r) - digamma(1), rank=row_number(), 
          p.adj= r*hm*p_value/rank)

proteins_2<-ttests_1 %>% 
  slice_min(p.adj, n =20) %>% 
  pull(protein)
```

```{r}
proteins_2 # top 20
```

Correlation with ADOS
```{r}
asd_clean <- biomarker_train %>% 
  select(group, ados) %>% 
  group_by(group) %>% 
  sample_n(size = 2)  # ados is only measured for the ASD group
asd_clean

biomarker_clean[76,]

biomarker_clean %>%
  group_by(group)

cors <- data.frame(matrix(ncol= 2, nrow=0))


ASD_proteins <-biomarker_clean[1:76,-c(1,2)]
# currently working on
for (i in 1:length(ASD_proteins)){
  #print(i)
  cor_i <- cor(biomarker_clean[1:76,]$ados, ASD_proteins[i]) 
  
  
  row_i <- cbind(colnames(biomarker_train[,-c(1,2)])[i], cor_i[[1,1]])
  cors <- rbind(cors, row_i)
}
colnames(cors) <- c('protein', 'correlation')

cors


# not right
for (i in length(colnames(biomarker_train[,-c(1,2)])[1:4])){
  #print(i)
  cor_i <- cor(biomarker_clean[1:76,]$ados, biomarker_clean[1:76,-c(1,2)][i])
  
  
  row_i <- cbind(colnames(biomarker_train[,-c(1,2)])[i],cor_i[[1,1]])
  cors <- rbind(cors, row_i)
}


cors

# compute correlations, not right
asd_clean %>%
  pivot_longer(cols = -ados,
               names_to = 'protein',
               values_to = 'level') %>%
  group_by(protein) %>%
  summarize(correlation = cor(ados, level))

biomarker_clean[,2:23]%>%
  pivot_longer(cols = -ados,
               names_to = 'protein',
               values_to = 'level') %>%
  group_by(protein) %>%
  summarize(correlation = cor(ados, level))
```

```{r}
proteins_2 # top 20
```

Logistic Regression 

```{r}
proteins_subset<-intersect(proteins_1,proteins_2)

biomarker_sset<- biomarker_train %>% 
  select(group, any_of(proteins_subset)) %>% 
  mutate(class= (group == 'ASD')) %>% 
  select(-group)

set.seed(549)
biomarker_split2<-biomarker_sset %>% 
  initial_split(prop=0.70)

fit<-glm(class ~., data=training(biomarker_split2), family= 'binomial')

class_metrics<-metric_set(sensitivity, specificity, accuracy,roc_auc)

testing(biomarker_split2) %>% 
  add_predictions(fit, type= 'response') %>% 
  class_metrics(estimate = factor(pred > 0.5), 
                truth= factor(class), pred, event_level = 'second')
```

top 10 
```{r}
fit %>% 
  broom::tidy() %>% 
  knitr::kable() 
```

```{r}
library(arsenal)

protein_1<-data.frame(proteins_1)
protein_2<-data.frame(proteins_2)

names(protein_1)[names(protein_1) == "proteins_1"] <- "protein"
names(protein_2)[names(protein_2) == "proteins_2"] <- "protein"
common_protein<-generics::intersect(protein_1,protein_2)
view(common_protein)
```

In a fuzzy intersection we examine the elements that are common in at least n of the
sets. Therefore in this experiment we can look at the proteins that were selected in
two of the models. In this case we used the random forest model and the t-test model
to examine which ten proteins were common among them. After using each model on our 
data we found that the "Top 10" common proteins were IgD, Cadherin-5, RELT, DERM, 
ILT-4, Troponin I, skeletal,fast twitch, MAPK2, MATN2, IGFBP-1,M2-PK. 